Does the leaf economic spectrum hold within plant functional types? A Bayesian multivariate trait meta-analysis

### 2019-11-22 shiklomanov_formatted.tex
Formatting revisions to formatted LaTeX

\usepackage{setspace}
\usepackage{fontspec}
\usepackage{graphicx}

@@

One such constraint is the "leaf economic spectrum", which defines a negative relationship between SLA and leaf lifespan, and a positive relationship of SLA with $N_{mass}$, $P_{mass}$, and photosynthesis and respiration rates \citep{wright_2004_worldwide,shipley_2006_fundamental,reich_2014_world,diaz_2015_global}.
Leaf economic traits are correlated with
plant productivity \citep{shipley_2005_functional,niinemets_2016_within,wu_2016_convergence},
litter decomposition rates \citep{bakker_2010_leaf,hobbie_2015_plant},
community composition \citep{burns_2004_patterns,cavender-bares_2004_multiple},
and ecosystem function \citep{diaz_2004_plant,musavi_2015_imprint}.
The position of species along the leaf economic spectrum is related to climate and soil conditions
\citep{wright_2004_worldwide,wright_2005_modulation,cornwell_2009_community,ordonez_2009_global,wigley_2016_leaf}.
As a result, relationships between leaf economic traits and climate have been used in ecosystem models to more finely resolve variation in plant function \citep{sakschewski_2015_leaf,verheijen_2015_inclusion,scheiter_2013_next}.

@@

An intermediate strategy is to aggregate species with similar structural and functional characteristics into plant functional types (PFTs).
Although PFTs are most widely used as the unit of plant functional differentiation in dynamic vegetation and earth system models
\citep{lavorel_1997_plant_functional_classifications,wullschleger_2014_plant,prentice_1992_special_paper},
the underlying concept of plant assemblages has been an important part of ecological discourse for over a century \citep{cowles_1899_ecological,clements_1936_nature,naeem_2003_disentangling}.
Patterns of trait variation within and across PFTs are relevant for several reasons.
First, trait covariance can be leveraged to impute missing trait values \citep{swenson_2013_phylogenetic};
the extent to which leveraging trait covariance reduces the size (i.e. variability and/or uncertainty) of the trait space has important implications for
quantifying the parametric uncertainty in vegetation model projections \citep{dietze_2013_improving,lebauer_2013_facilitating,dietze_2014_quantitative}.
Second, the large uncertainty in model projections of future global carbon budgets \citep{friedlingstein_2006_climate,friedlingstein_2014_uncertainties}
has been at least partially attributed to the models' failure to account for plant adaptation to changing environments \citep{sitch_2008_evaluation},
which has led to an increased interest in adding within-PFT trait plasticity to models \citep{van_2011_going,verheijen_2015_inclusion}.

While the leaf economic spectrum has been investigated at the global scale, where it is robust, and at local scales, where deviations from it are common, it has received less attention at the intermediate scale of PFTs.
Thus, this paper seeks to answer the following questions:
First, does the leaf economic spectrum hold within vs. across PFTs?
Second, can the leaf economic spectrum and similar covariance patterns be leveraged to reduce uncertainties in trait estimates, particularly under data limitation?
The answers to these question have implications for both functional ecology and ecosystem modeling.
To address these questions, we developed a hierarchical multivariate Bayesian model that explicitly accounts for across- and within-PFT variability in trait correlations.
We then fit this model to a global trait database to estimate mean trait values and variance-covariance matrices for PFTs as defined in a major earth system model (Community Land Model, CLM, \citealt{clm45_note}).
We evaluate the ability of this model to reduce uncertainties in trait estimates and reproduce observed patterns of global trait variation compared to univariate models.
Finally, we assess the scale dependence and generality of estimated trait covariances.

@@

We focused on seven leaf traits obtained from the TRY global database \citep{kattge_2011_try} (see Appendix S1):
longevity (months),
specific leaf area (SLA, m$^2$ kg$^{-1}$),
nitrogen content ($N_{mass}$, mg N g$^{-1}$ or $N_{area}$, g m$^{-2}$),
phosphorus content ($P_{mass}$, mg P g$^{-1}$ or $P_{area}$, g m$^{-2}$),
dark respiration at 25°C ($R_{d,mass}$, µmol g$^{-1}$ s$^{-1}$, or $R_{d,area}$, µmol m$^{-2}$ s$^{-1}$),
maximum RuBisCO carboxylation rate at 25°C ($V_{c,\max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $V_{c,\max,area}$, µmol m$^{-2}$ s$^{-1}$),
and maximum electron transport rate at 25°C ($J_{\max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{\max,area}$, µmol m$^{-2}$ s$^{-1}$).
For $V_{c,\max}$, we only used values reported at 25°C.
For $R_{d}$ and $J_{\max}$, we normalized the values to 25°C using reported leaf temperature values following \citet{atkin_2015_global} and \citet{kattge_2007_temperature} (equation 1 therein), respectively.
To avoid issues with trait normalization, we performed analyses separately for both mass- and area-normalized traits \citep{osnas_2013_global,lloyd_2013_photosynthetically}.
We restricted our analysis to quality-controlled values from species with sufficient information for functional type classification \cite[see][]{kattge_2011_try}.
Following past studies \cite[e.g.][]{wright_2004_worldwide,onoda_2011_global,diaz_2015_global}, we log-transformed all trait values to correct for their strong right-skewness.

@@

We assigned each species a PFT following the scheme in the Community Land Model (CLM4.5, \citealt{clm45_note}) (Tab. 1, Fig. 1).
We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from TRY.
Where species attributes disagreed between datasets, we assigned the most frequently observed attribute (e.g., if five datasets say "shrub" but only one says "tree", we would use "shrub").
Where species attributes were missing, we assigned attributes based on higher order phylogeny if possible (e.g., /Poaceae/ family are grasses, /Larix/ genus are deciduous needleleaved trees) or omitted the species if not.
For biome specification, we matched geographic coordinates for each species to annual mean temperature ($AMT$, averaged 1970-2000) data from WorldClim-2 \citep{fick_2017_worldclim},
calculated the mean $AMT$ for all sites where each species was observed,
and then binned these species based on the following cutoffs: boreal/arctic ($AMT \leq 5^\circ C$), temperate ($AMT \leq 20^\circ C$), and tropical ($AMT > 20^\circ C$).

@@

We compared three models with different levels of complexity.
The simplest was the "univariate" model, in which each trait is independent.
For an observation $x_{i,t}$ of trait $t$ and sample $i$:

@@

The second-simplest model was the "multivariate" model, in which traits are drawn from a single multivariate distribution.
For observed trait vector ${\mathbf{x_i}}$ for sample $i$:

\begin{equation}
\mathbf{x_i} \sim mvN(\mathbf{\mu}, \mathbf{\Sigma})
\end{equation}

where $mvN$ is the multivariate Gaussian distribution with mean vector $\mathbf{\mu}$ and covariance matrix $\mathbf{\Sigma}$.
We fit both of these models independently for each PFT and once for the entire dataset (i.e. one global PFT).

The most complex model was the ``hierarchical multivariate'' model (henceforth, just ``hierarchical model''),
where traits are drawn from a PFT-specific multivariate distribution describing within-PFT variation,
and whose mean vector is itself sampled from a global multivariate distribution describing variation across PFTs.
For observed trait vector $\mathbf{x}_{i,p}$ for sample $i$ belonging to PFT $p$:

\begin{equation}
\mathbf{x}_{i,p} \sim mvN(\mathbf{\mu}_p, \mathbf{\Sigma}_p)
\end{equation}

\begin{equation}
\mathbf{\mu}_p \sim mvN(\mathbf{\mu}_g, \mathbf{\Sigma}_g)
\end{equation}

where $\mathbf{\mu}_p$ and $\mathbf{\Sigma}_p$ are the mean vector and covariance matrix describing variation within PFT $p$, and $\mathbf{\mu}_g$ and $\mathbf{\Sigma}_g$ are the mean vector and covariance matrix describing across-PFT (global) variation.

@@

For priors on all multivariate mean vectors ($\mathbf{\mu}$), we used multivariate normal distributions.
For priors on all multivariate variance-covariance matrices, we used the Wishart distribution ($W$), which leads to the following posterior distribution:

\begin{equation}
P(\mathbf{\Sigma} \mid
\mathbf{x}, \mathbf{\mu},
\nu_0, \mathbf{\Sigma}_0)
\sim
(W(\nu^*, S^*))^{-1}
\end{equation}

@@

\begin{equation}
\mathbf{S^*} = (\mathbf{S}_0 + (\bar{\mathbf{x}} - \mu)^T (\bar{\mathbf{x}} - \mu))^{-1}
\end{equation}

where $n$ is the number of observations, $m$ is the number of traits in data matrix $\mathbf{x}$, and $\bar{\mathbf{x}}$ is the column means of $\mathbf{x}$.
For further details on the derivation of the conjugate relationship, see \citet[Section 3.6, "Multivariate normal with unknown mean and variance", pg. 72]{gelman_2003_bayesian}.

We used weakly-informative priors for trait means and variances (diagonals of the multivariate normal covariance matrix), the values of which are shown in Appendix S2: Table S1.
All of the covariance (off-diagonal) terms in the prior variance matrix were set to zero.
We used uninformative priors for the Wishart distribution ($\nu_0 = 0$, $\mathbf{S}_0 = \mathrm{diag}(1, m)$).

The above equations defining the conjugacy relationship do not work if the data matrix $x$ has any missing values.
Therefore, we modeled the partially missing observations as latent variables conditioned on the present observations and estimated mean vector and covariance matrix.
This approach is conceptually similar to multiple imputation \citep{white_2010_multiple,graham_2009_missing_data_analysis},
and is quite distinct from single imputation, where data are imputed once in a separate step prior to parameter estimation \citep{white_2010_multiple,graham_2009_missing_data_analysis}.
For a block of data $\mathbf{x\prime}$ containing missing observations in columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$,
missing values $\mathbf{x\prime}[m]$ are drawn randomly from a conditional multivariate normal distribution at each iteration of the sampling algorithm:

\begin{equation}
\mathbf{x^\prime}[m|p] \sim mvN(\mathbf{\mu}^\prime, \mathbf{\Sigma}^\prime)
\end{equation}

\begin{equation}
\mathbf{\mu\prime} =
(\mathbf{x\prime}[p] - \mathbf{\mu^\prime}[p])
(\mathbf{\Sigma}[p,p]^{-1} \mathbf{\Sigma}[p,m])
\end{equation}

\begin{equation}
\mathbf{\Sigma\prime} = \mathbf{\Sigma}[m,m] -
\mathbf{\Sigma}[m,p]
(\mathbf{\Sigma}[p,p]^{-1} \mathbf{\Sigma}[p,m])
\end{equation}

Sampling proceeds according to the following algorithm:
Let $\mu_i$ and $\mathbf{\Sigma}_i$ be the estimates of the mean vector and covariance matrix, respectively, at MCMC iteration $i$.
Similarly, let $x_i$ be the realization of the data $\mathbf{x}\prime$ with missing (latent) values imputed at MCMC iteration $i$.

1. Initialize $\mu_1$ and $\mathbf{\Sigma}_1$ as a random draw from their respective priors.
2. Generate $x_1$ as a function of $\mu_1$ and $\mathbf{\Sigma}_1$.
3. Draw $\mu_2$ and $\mathbf{Sigma}_2$ as a function of $x_1$, according to the corresponding Gibbs sampling step.
4. Generate $x_2$ as function of $\mu_2$ and $\mathbf{\Sigma}_2$.
5. Draw $\mu_3$ and $\mathbf{Sigma}_3$ as a function of $x_2$.
6. Continue alternating these steps until a stable distribution of $\mu$ and $\mathbf{Sigma}$ is reached.

@@

To assess the consistency of within- and across-PFT trait trade-offs,
we calculated the mean and 95\% credible interval of the pairwise reduced major axis slope ($M$) for each trait pair ($i$, $j$)
from posterior samples of their variance-covariance matrices ($\mathbf{\Sigma}$) using the following equation:

\begin{equation}
M_{i,j} = \frac{\mathbf{\Sigma}_{j,j}}{\mathbf{\Sigma}_{i,i}} \textrm{sign}(\mathbf{\Sigma}_{i,j})
\end{equation}

@@

All R analyses were run using R version 3.6.1 \citep{r_361}.
The R code and data for running these analyses is publicly available online via the Open Science Framework at https://osf.io/w8y73/.
To comply with TRY intellectual property guidelines, the trait data used in this study have been "anonymized" such that they can only be identified to the PFT level (not the species level) as required to reproduce this analysis.
The complete TRY data request used for this analysis has been archived at http://try-db.org, and can be retrieved by providing the TRY data request ID (1584).

@@

Morphological and biochemical traits (SLA, N, P) showed the most robust and consistent mutual covariance of these three groups.
SLA was positively related to $N_{mass}$ and $P_{mass},$ and negatively related to $N_{area}$ and $P_{area}$, both across PFTs and within all PFTs.
The magnitude of the slopes between N and P (regardless of normalization), and of SLA with $N_{area}$ and $P_{area}$, were relatively constant within all PFTs,
but the magnitude of the slopes of SLA with $N_{mass}$ and $P_{mass}$ were more variable.
In particular, temperate tree species (BlETe, BlDTe, NlETe) showed steeper SLA-$N_{mass}$ slopes (more variation in SLA relative to $N_{mass}$) than most other PFTs.

@@

An important caveat to these results is that many slopes, including all of the across-PFT slopes, had 95\% credible intervals that intersected zero---i.e. we are less than 95\% confident in the direction of these slopes.
This is primarily due to variations in the effective number of pairwise observations used to estimate the covariance matrix ---
the more pairwise observations are available, the smaller the minimum covariance that can be estimated with the same level of statistical power and confidence.
For example, a power analysis of correlation coefficients (`pwr::pwr.r.test` in R; \citealt{r_pwr_package}) showed that with 14 plant functional types ($n = 14$),
the smallest across-PFT correlation we would be able to estimate with 95\% power ($\alpha = 0.95$) and confidence ($p = 0.05$) is 0.74,
so we can confidently say that all PFT correlation coefficients (different from, but closely related to slope) were smaller than that value.
That being said, because all across-PFT slopes have the same sample size,
we can reasonably expect differences in the mean strength of pairwise across-PFT trait relationships to be ecologically meaningful.
The situation is more complex for PFT-level estimates, where sample size varies by multiple orders of magnitude by PFT and trait pair (Tab. 1, Appendix S2: Tab. S4).
In particular, high-latitude PFTs (BlDBo, NlEBo, NlD, ShDBo, and C3GAr) and metabolic traits ($R_d$, $V_{c,\max}$, $J_{\max}$) stand out as having particularly low sample sizes.

>>>>

\usepackage{setspace}
\usepackage{fontspec}
\usepackage{graphicx}
\usepackage{bm}

@@

One such constraint is the ``leaf economic spectrum'', which defines a negative relationship between SLA and leaf lifespan, and a positive relationship of SLA with $N_{mass}$, $P_{mass}$, and photosynthesis and respiration rates~\citep{wright_2004_worldwide,shipley_2006_fundamental,reich_2014_world,diaz_2015_global}.
Leaf economic traits are correlated with
plant productivity \citep{shipley_2005_functional,niinemets_2016_within,wu_2016_convergence},
litter decomposition rates \citep{bakker_2010_leaf,hobbie_2015_plant},
community composition \citep{burns_2004_patterns,cavender-bares_2004_multiple},
and ecosystem function \citep{diaz_2004_plant,musavi_2015_imprint}.
The position of species along the leaf economic spectrum is related to climate and soil conditions
\citep{wright_2004_worldwide,wright_2005_modulation,cornwell_2009_community,ordonez_2009_global,wigley_2016_leaf}.
As a result, relationships between leaf economic traits and climate have been used in ecosystem models to more finely resolve variation in plant function \citep{sakschewski_2015_leaf,verheijen_2015_inclusion,scheiter_2013_next}.

@@

An intermediate strategy is to aggregate species with similar structural and functional characteristics into plant functional types (PFTs).
Although PFTs are most widely used as the unit of plant functional differentiation in dynamic vegetation and earth system models
\citep{lavorel_1997_plant_functional_classifications,wullschleger_2014_plant,prentice_1992_special_paper},
the underlying concept of plant assemblages has been an important part of ecological discourse for over a century \citep{cowles_1899_ecological,clements_1936_nature,naeem_2003_disentangling}.
Patterns of trait variation within and across PFTs are relevant for several reasons.
First, trait covariance can be leveraged to impute missing trait values \citep{swenson_2013_phylogenetic};
the extent to which leveraging trait covariance reduces the size (i.e.\ variability and/or uncertainty) of the trait space has important implications for
quantifying the parametric uncertainty in vegetation model projections \citep{dietze_2013_improving,lebauer_2013_facilitating,dietze_2014_quantitative}.
Second, the large uncertainty in model projections of future global carbon budgets \citep{friedlingstein_2006_climate,friedlingstein_2014_uncertainties}
has been at least partially attributed to the models' failure to account for plant adaptation to changing environments \citep{sitch_2008_evaluation},
which has led to an increased interest in adding within-PFT trait plasticity to models \citep{van_2011_going,verheijen_2015_inclusion}.

While the leaf economic spectrum has been investigated at the global scale, where it is robust, and at local scales, where deviations from it are common, it has received less attention at the intermediate scale of PFTs.
Thus, this paper seeks to answer the following questions:
First, does the leaf economic spectrum hold within vs.\ across PFTs?
Second, can the leaf economic spectrum and similar covariance patterns be leveraged to reduce uncertainties in trait estimates, particularly under data limitation?
The answers to these question have implications for both functional ecology and ecosystem modeling.
To address these questions, we developed a hierarchical multivariate Bayesian model that explicitly accounts for across- and within-PFT variability in trait correlations.
We then fit this model to a global trait database to estimate mean trait values and variance-covariance matrices for PFTs as defined in a major earth system model (Community Land Model, CLM, \citealt{clm45_note}).
We evaluate the ability of this model to reduce uncertainties in trait estimates and reproduce observed patterns of global trait variation compared to univariate models.
Finally, we assess the scale dependence and generality of estimated trait covariances.

@@

We focused on seven leaf traits obtained from the TRY global database \citep{kattge_2011_try} (see Appendix S1):
longevity (months),
specific leaf area (SLA, m$^2$ kg$^{-1}$),
nitrogen content ($N_{mass}$, mg N g$^{-1}$ or $N_{area}$, g m$^{-2}$),
phosphorus content ($P_{mass}$, mg P g$^{-1}$ or $P_{area}$, g m$^{-2}$),
dark respiration at 25°C ($R_{d,mass}$, µmol g$^{-1}$ s$^{-1}$, or $R_{d,area}$, µmol m$^{-2}$ s$^{-1}$),
maximum RuBisCO carboxylation rate at 25°C ($V_{c,\max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $V_{c,\max,area}$, µmol m$^{-2}$ s$^{-1}$),
and maximum electron transport rate at 25°C ($J_{\max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{\max,area}$, µmol m$^{-2}$ s$^{-1}$).
For $V_{c,\max}$, we only used values reported at 25°C.
For $R_{d}$ and $J_{\max}$, we normalized the values to 25°C using reported leaf temperature values following \citet{atkin_2015_global} and \citet{kattge_2007_temperature} (equation 1 therein), respectively.
To avoid issues with trait normalization, we performed analyses separately for both mass- and area-normalized traits \citep{osnas_2013_global,lloyd_2013_photosynthetically}.
We restricted our analysis to quality-controlled values from species with sufficient information for functional type classification~\cite[see][]{kattge_2011_try}.
Following past studies~\cite[e.g.][]{wright_2004_worldwide,onoda_2011_global,diaz_2015_global}, we log-transformed all trait values to correct for their strong right-skewness.

@@

We assigned each species a PFT following the scheme in the Community Land Model (CLM4.5, \citealt{clm45_note}) (Tab. 1, Fig. 1).
We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from TRY.
Where species attributes disagreed between datasets, we assigned the most frequently observed attribute (e.g., if five datasets say ``shrub'' but only one says ``tree'', we would use ``shrub'').
Where species attributes were missing, we assigned attributes based on higher order phylogeny if possible (e.g., \textit{Poaceae} family are grasses, \textit{Larix} genus are deciduous needleleaved trees) or omitted the species if not.
For biome specification, we matched geographic coordinates for each species to annual mean temperature ($AMT$, averaged 1970--2000) data from WorldClim-2 \citep{fick_2017_worldclim},
calculated the mean $AMT$ for all sites where each species was observed,
and then binned these species based on the following cutoffs: boreal/arctic ($AMT \leq 5^\circ C$), temperate ($AMT \leq 20^\circ C$), and tropical ($AMT > 20^\circ C$).

@@

We compared three models with different levels of complexity.
The simplest was the ``univariate'' model, in which each trait is independent.
For an observation $x_{i,t}$ of trait $t$ and sample $i$:

@@

The second-simplest model was the ``multivariate'' model, in which traits are drawn from a single multivariate distribution.
For observed trait vector ${\bm{x_i}}$ for sample $i$:

\begin{equation}
\bm{x_i} \sim mvN(\bm{\mu}, \bm{\Sigma})
\end{equation}

where $mvN$ is the multivariate Gaussian distribution with mean vector $\bm{\mu}$ and covariance matrix $\bm{\Sigma}$.
We fit both of these models independently for each PFT and once for the entire dataset (i.e.\ one global PFT).

The most complex model was the ``hierarchical multivariate'' model (henceforth, just ``hierarchical model''),
where traits are drawn from a PFT-specific multivariate distribution describing within-PFT variation,
and whose mean vector is itself sampled from a global multivariate distribution describing variation across PFTs.
For observed trait vector $\bm{x}_{i,p}$ for sample $i$ belonging to PFT $p$:

\begin{equation}
\bm{x}_{i,p} \sim mvN(\bm{\mu}_p, \bm{\Sigma}_p)
\end{equation}

\begin{equation}
\bm{\mu}_p \sim mvN(\bm{\mu}_g, \bm{\Sigma}_g)
\end{equation}

where $\bm{\mu}_p$ and $\bm{\Sigma}_p$ are the mean vector and covariance matrix describing variation within PFT $p$, and $\bm{\mu}_g$ and $\bm{\Sigma}_g$ are the mean vector and covariance matrix describing across-PFT (global) variation.

@@

For priors on all multivariate mean vectors ($\bm{\mu}$), we used multivariate normal distributions.
For priors on all multivariate variance-covariance matrices, we used the Wishart distribution ($W$), which leads to the following posterior distribution:

\begin{equation}
P(\bm{\Sigma} \mid
\bm{x}, \bm{\mu},
\nu_0, \bm{\Sigma}_0)
\sim
(W(\nu^*, S^*))^{-1}
\end{equation}

@@

\begin{equation}
\bm{S^*} = (\bm{S}_0 + (\bar{\bm{x}} - \mu)^T (\bar{\bm{x}} - \mu))^{-1}
\end{equation}

where $n$ is the number of observations, $m$ is the number of traits in data matrix $\bm{x}$, and $\bar{\bm{x}}$ is the column means of $\bm{x}$.
For further details on the derivation of the conjugate relationship, see \citet[Section 3.6, "Multivariate normal with unknown mean and variance", pg. 72]{gelman_2003_bayesian}.

We used weakly-informative priors for trait means and variances (diagonals of the multivariate normal covariance matrix), the values of which are shown in Appendix S2: Table S1.
All of the covariance (off-diagonal) terms in the prior variance matrix were set to zero.
We used uninformative priors for the Wishart distribution ($\nu_0 = 0$, $\bm{S}_0 = \mathrm{diag}(1, m)$).

The above equations defining the conjugacy relationship do not work if the data matrix $x$ has any missing values.
Therefore, we modeled the partially missing observations as latent variables conditioned on the present observations and estimated mean vector and covariance matrix.
This approach is conceptually similar to multiple imputation \citep{white_2010_multiple,graham_2009_missing_data_analysis},
and is quite distinct from single imputation, where data are imputed once in a separate step prior to parameter estimation \citep{white_2010_multiple,graham_2009_missing_data_analysis}.
For a block of data $\bm{x\prime}$ containing missing observations in columns $\bm{m}$ and present observations in columns $\bm{p}$,
missing values $\bm{x\prime}[m]$ are drawn randomly from a conditional multivariate normal distribution at each iteration of the sampling algorithm:

\begin{equation}
\bm{x^\prime}[m|p] \sim mvN(\bm{\mu}^\prime, \bm{\Sigma}^\prime)
\end{equation}

\begin{equation}
\bm{\mu\prime} =
(\bm{x\prime}[p] - \bm{\mu^\prime}[p])
(\bm{\Sigma}[p,p]^{-1} \bm{\Sigma}[p,m])
\end{equation}

\begin{equation}
\bm{\Sigma\prime} = \bm{\Sigma}[m,m] -
\bm{\Sigma}[m,p]
(\bm{\Sigma}[p,p]^{-1} \bm{\Sigma}[p,m])
\end{equation}

Sampling proceeds according to the following algorithm:
Let $\mu_i$ and $\bm{\Sigma}_i$ be the estimates of the mean vector and covariance matrix, respectively, at MCMC iteration $i$.
Similarly, let $x_i$ be the realization of the data $\bm{x}\prime$ with missing (latent) values imputed at MCMC iteration $i$.

\begin{enumerate}
\item Initialize $\mu_1$ and $\bm{\Sigma}_1$ as a random draw from their respective priors.
\item Generate $x_1$ as a function of $\mu_1$ and $\bm{\Sigma}_1$.
\item Draw $\mu_2$ and $\bm{\Sigma}_2$ as a function of $x_1$, according to the corresponding Gibbs sampling step.
\item Generate $x_2$ as function of $\mu_2$ and $\bm{\Sigma}_2$.
\item Draw $\mu_3$ and $\bm{\Sigma}_3$ as a function of $x_2$.
\item Continue alternating these steps until a stable distribution of $\mu$ and $\bm{\Sigma}$ is reached.
\end{enumerate}

@@

To assess the consistency of within- and across-PFT trait trade-offs,
we calculated the mean and 95\% credible interval of the pairwise reduced major axis slope ($M$) for each trait pair ($i$, $j$)
from posterior samples of their variance-covariance matrices ($\bm{\Sigma}$) using the following equation:

\begin{equation}
M_{i,j} = \frac{\bm{\Sigma}_{j,j}}{\bm{\Sigma}_{i,i}} \textrm{sign}(\bm{\Sigma}_{i,j})
\end{equation}

@@

All R analyses were run using R version 3.6.1 \citep{r_361}.
The R code and data for running these analyses is publicly available online via the Open Science Framework at https://osf.io/w8y73/.
To comply with TRY intellectual property guidelines, the trait data used in this study have been ``anonymized'' such that they can only be identified to the PFT level (not the species level) as required to reproduce this analysis.
The complete TRY data request used for this analysis has been archived at http://try-db.org, and can be retrieved by providing the TRY data request ID (1584).

@@

Morphological and biochemical traits (SLA, N, P) showed the most robust and consistent mutual covariance of these three groups.
SLA was positively related to $N_{mass}$ and $P_{mass}$, and negatively related to $N_{area}$ and $P_{area}$, both across PFTs and within all PFTs.
The magnitude of the slopes between N and P (regardless of normalization), and of SLA with $N_{area}$ and $P_{area}$, were relatively constant within all PFTs,
but the magnitude of the slopes of SLA with $N_{mass}$ and $P_{mass}$ were more variable.
In particular, temperate tree species (BlETe, BlDTe, NlETe) showed steeper SLA-$N_{mass}$ slopes (more variation in SLA relative to $N_{mass}$) than most other PFTs.

@@

An important caveat to these results is that many slopes, including all of the across-PFT slopes, had 95\% credible intervals that intersected zero---i.e.\ we are less than 95\% confident in the direction of these slopes.
This is primarily due to variations in the effective number of pairwise observations used to estimate the covariance matrix ---
the more pairwise observations are available, the smaller the minimum covariance that can be estimated with the same level of statistical power and confidence.
For example, a power analysis of correlation coefficients (`pwr::pwr.r.test' in R;\@\citealt{r_pwr_package}) showed that with 14 plant functional types ($n = 14$),
the smallest across-PFT correlation we would be able to estimate with 95\% power ($\alpha = 0.95$) and confidence ($p = 0.05$) is 0.74,
so we can confidently say that all PFT correlation coefficients (different from, but closely related to slope) were smaller than that value.
That being said, because all across-PFT slopes have the same sample size,
we can reasonably expect differences in the mean strength of pairwise across-PFT trait relationships to be ecologically meaningful.
The situation is more complex for PFT-level estimates, where sample size varies by multiple orders of magnitude by PFT and trait pair (Tab. 1, Appendix S2: Tab. S4).
In particular, high-latitude PFTs (BlDBo, NlEBo, NlD, ShDBo, and C3GAr) and metabolic traits ($R_d$, $V_{c,\max}$, $J_{\max}$) stand out as having particularly low sample sizes.
